Analyses of 27 years of stand structure development and basal area growth response to a range of initial basal area density levels, 1992-2019

### 2020-04-30 manuscript/EP1162_manuscript_draft.Rmd
starting on manuscript

---
title: "Tree Level Changes"
author: "Hardy Griesbauer"
date: "21/04/2020"
output: html_document
---

@@

```
Below is code and output looking at tree-level changes over time

Also, we can include plots 16-19 even though they weren't measured across all periods, because we are looking at tree-level responses.

### PBAI at the tree level
We can look at PBAi at the tree-level, by tree class.  This will tell us where PBAI is concentrated.  Hypothesis is that smaller trees respond with higher PBAI and responses are stronger with in plotss with higher basal area removal.

```{r treeList and other starting variables}

# Let's filter only for trees that have five measurements 
treeList<- # assign variable
  dat %>% 
  filter(TreeID!="New") %>% #remove new trees
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(Species %in% c("Sx","Bl")) %>% # filter birch out of dataset
  filter(Status=="Live") %>% # filter for live trees
  group_by(TreeID) %>% 
  summarise(N=n()) %>% 
  filter(N==5) %>% #filter for trees with 5 measurements
  pull(TreeID)

plotBA<-
  dat %>% 
  filter(Status=="Live") %>% # filter for live trees
   mutate(BA=(DBH/2/100)^2*pi) %>% # convert DBH to basal area 
  group_by(Plot,Year) %>% # grouping
  summarise(plotBA=sum(BA))

# derive PBAI
pbaiDat<-
  dat %>% 
  mutate(TreeID=paste(Plot,TreeID,sep="-")) %>% # create unique treeID
  filter(TreeID %in% treeList) %>% 
  mutate(BA=(DBH/2/100)^2*pi) %>% # convert DBH to basal area
  dplyr::select(Plot,TreeID,Species,Year,BA) %>% 
  pivot_wider(names_from=Year,values_from=BA) %>% 
  mutate(PBAI=`2019`-`1992`) %>% 
  dplyr::select(Plot,TreeID,Species,PBAI,baInit=`1992`)

@@

LMER to predict basal area slope per tree after treatment
```{r}
x<- # assign output 
  pbaiDat %>% 
  left_join(plotDat,by="Plot")

```{r lmer by PBAI}
fit1<-
  x %>% 
  lmer(log(PBAI)~BA.Target+Species+log(baInit)+(1|Plot),data=.)

Anova(fit1)

plot(fit1)

glance(fit1)

```{r plot marginal means}

mydf<-  ggpredict(fit1,terms=c("baInit [exp]","Species","BA.Target"))

ggplot(mydf, aes(x, predicted, colour = group)) + 
  geom_line() +
  facet_wrap(~facet)+
  ylab("Predicted BAI (m2) per tree from 1992 to 2019")+
  xlab("Tree basal area (m2) in 1992")

ggsave("predictedPBAI.png")
```

```{r r squared values from model}

r.squaredGLMM(fit1)

## Let's look at basal area composition by tree size per treatment per year
```{r ba by tree size, year and treatment}

# set up diameter classes
diamClass=seq(from=7.5,to=95,by=5)

dat %>% 
   filter(!Plot %in% c(16,17,18,19)) %>%  # remove plots 16-19 from data for further analysis
  left_join(plotDat,by="Plot") %>% # join with plot data
   filter(Status=="Live") %>% # filter for live trees only
   mutate(DBH.class=cut(DBH,diamClass,ordered_result = TRUE,right=F)) %>% # assign DBH classes
  mutate(BA=(DBH/2/100)^2*pi) %>% # convert DBH to basal area
  group_by(BA.Target,Plot,Year,DBH.class) %>% 
  summarise(BA.Plot=sum(BA)*20) %>%   # sum basal area per diam class and multiply to get per ha

  # create means
  ungroup() %>% 
  group_by(BA.Target,Year,DBH.class) %>%
  summarise(BA=mean(BA.Plot)) %>%

   # Add labeller column
  mutate(label=factor(paste("Trt:",BA.Target,sep=""))) %>%

  # Plotting
  ggplot()+
  aes(x=DBH.class,y=BA,fill=label)+
  geom_bar(stat="Identity")+
  facet_grid(Year~label)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,size=6,vjust=0.25))+
  ylab("Mean basal area per hectare per plot (m2)")+
  xlab("Diameter classes (cm)")

ggsave(filename="baDiamClass.png")

>>>>

---
title: Analyses of 27 years of stand structure development and basal area growth response to a range of initial basal area density levels, 1992-2019 
author: "Mike Jull and Hardy Griesbauer"
date: "29/03/2020"
output:
  word_document:
    reference_docx: wordTemplate2.docx
  html_document:
    df_print: paged
csl: cjfr.csl
bibliography: ep1162Bib.bib
---

@@

@@
